### 2017-01-09 local-capability-machine.tex
scall lemma

% Variables
\newcommand{\var}[1]{\mathit{#1}}
\newcommand{\hs}{\var{ms}}
\newcommand{\ms}{\hs}
\newcommand{\hv}{\var{hv}}
\newcommand{\rv}{\var{rv}}
\newcommand{\lv}{\var{lv}}
\newcommand{\gl}{\var{g}}
\newcommand{\pc}{\mathit{pc}}
\newcommand{\pcreg}{\mathrm{pc}}
\newcommand{\addr}{\var{a}}
\newcommand{\offset}{\var{offset}}
\newcommand{\word}{\var{w}}
\newcommand{\start}{\var{base}}
\newcommand{\addrend}{\var{end}}
\newcommand{\pwlv}{\var{pwl}}
\newcommand{\mem}{\var{mem}}
\newcommand{\reg}{\var{reg}}
\newcommand{\heapseg}{\var{ms}}
\newcommand{\heap}{\var{mem}}
\newcommand{\mode}{\var{mode}}
\newcommand{\perm}{\var{perm}}
\newcommand{\permp}{\var{permPair}}
\newcommand{\roll}{\var{roll}}
\newcommand{\instr}{\var{instr}}
\newcommand{\stdcap}[1][(\perm,\gl)]{\left(#1,\start,\addrend,\addr \right)}
\newcommand{\adv}{\var{adv}}
\newcommand{\msframe}{ms_\var{frame}}
\newcommand{\link}{\var{link}}
\newcommand{\stk}{\var{stk}}
\newcommand{\flag}{\var{flag}}
\newcommand{\nwl}{\var{nwl}}
\newcommand{\pwl}{\var{pwl}}
\newcommand{\sta}{\var{sta}}
\newcommand{\cnst}{\var{cnst}}
\newcommand{\olf}{\var{offsetLinkFlag}}
\newcommand{\prp}{\var{prp}}

@@

% Macros
\newcommand{\scall}[3]{\mathtt{scall} \; #1(#2,#3)}

@@

\begin{proof}[Proof of Lemma~\ref{lem:correctness-f3}]
\lau{proof using scall lemma}
Assume the premises of the lemma.
Now define 
  \begin{align*}
    W = \;& [0 \mapsto \iota_{\malloc,0}] \\
          & [1 \mapsto \iota^\sta (\perma,\ms_\flag \uplus \ms_{f2})]\\
          & [2 \mapsto \iota^\sta (\perma,\ms_\link)]\\
          & [3 \mapsto \iota^{\nwl,p}_{\start_\adv,\addrend_\adv}]\\
  \end{align*}
Further define
\[
\ms' = \ms_\flag \uplus \ms_{f2} \uplus \ms_\link \uplus \ms_\adv
\]
If we can show 
\begin{equation}
\label{eq:f3-obs1}
  \npair[n+1]{(\reg\update{\pcreg}{c_{f3}}\update{r_\stk}{c_\stk},\ms_{\malloc} \uplus \ms' \uplus \ms_{\adv} \uplus \ms_\stk)} \in \observations(W),
\end{equation}
then using $\ms_{\var{frame}}$ as the frame and $m'$ as the resulting memory, we get that $m' = \ms'' \uplus \ms_r \uplus \ms_{\var{frame}}$ for some $\ms'$ and $\ms_r$ s.t. $\heapSat[1]{\ms''}{W}$. Region $1$ guarantees that the assertion flag is unchanged, so we have
\[
  m'(\flag) = 0
\]
So SFTS \ref{eq:f3-obs1}. To do so, we use Lemma \ref{lem:anti-red-obs}. Let $\ms_f$ be given. The execution proceeds as follows:
\[
  (\reg\update{\pcreg}{c_{f3}}\update{r_\stk}{c_\stk},\ms' \uplus \ms_\stk \uplus \ms_f) \step[i] (\reg',\ms' \uplus [\start_\stk \mapsto 1] \uplus \ms_\stk|_{\start_\stk+1,\addrend_\stk} \uplus \ms_f),
\]
where
\[
  \lookingat{(\reg',\ms')}{\scall{r}{[]}{[]}}{c_{\var{next}}}
\]
where $c_{\var{next}}$ is $c_{f3}$ adjusted to point to the next instruction, namely \texttt{pop r1}. Further we have \pointstostack{\reg'}{[\start_\stk \mapsto 1]}{\ms_\stk |_{\start_\stk+1,\addrend_\stk}} and $i$ is a suitable number of steps.

@@

            \npair[n']{(\reg_1\update{\pcreg}{\updatePcPerm{c_{g1}}},\ms_1} \in \observations(W_1)

@@

\begin{lemma}[$\mathtt{scall}$ works]
  \label{lem:scall-works}
  If
  \begin{itemize}
  \item $\heapSat[\ms]{n}{W}$ 
  \item $(\reg,\ms) \text{ is looking at }
    \mathtt{scall}(r(\overline{r_{\mathit{arg}}},
    \overline{r_{\mathit{priv}}})) \text{ followed by } c_{\mathit{next}}$
  \item $\reg \text{ points to stack with $\ms_\stk$ used and $\ms_{\mathit{unused}}$ unused}$
  \item[Hyp-Callee] If
    \begin{itemize}
    \item $\dom(\ms_{\mathit{unused}}) = \dom(\ms_{\mathit{act}}
      \uplus \ms_{\mathit{unused}}')$,
    \item $W' =
      \revokeTemp{W}[\iota^{\sta}(\temp,\ms_\stk\uplus\ms_{\mathit{act}}),\iota^{\pwl}(\dom(\ms_{\mathit{unused}}'))]$,
    \item $\memSat[n-1]{\ms''}{W'}$
    \item $\reg' \text{ points to stack with $\emptyset$ used and $\ms_{\mathit{unused}'}$ unused}$
    \item $\forall a \in \dom(\ms_{\var{unused}}') \ldotp \ms''(a) = 0$
    \item $\reg'= \reg_0[\pcreg\mapsto\updatePcPerm{\reg(r)},
      \overline{r_{\mathit{arg}}} \mapsto \reg(\overline{r_{\mathit{arg}}}),r_0
      \mapsto c_{\mathit{ret}}, r_\stk \mapsto \_, r \mapsto \reg(r)]$ 
    \item $(n,c_{\mathit{ret}}) \in \stdvr(W')$
    \end{itemize}

@@

    then we have that $(n', (\reg',\ms'' \uplus \ms_\stk \uplus \ms_{\var{unused}}'')) \in \observations(\revokeTemp{W''})$
  \end{itemize}
  Then 
  \begin{itemize}
    \item $(\reg,\ms \uplus \ms_\stk \uplus \ms_{\var{unused}}) \in \observations(W)$
  \end{itemize}

@@

\begin{lemma}[Anti-reduction for $\observations$]
\label{lem:anti-red-obs}
  \begin{align*}
    & \forall n,i,\reg, \reg', \ms, \ms', \ms_r, W \ldotp\\
    & \quad (\forall \ms_f \ldotp (\reg,\ms \uplus \ms_r \uplus \ms_f) \step[i] (\reg',\ms' \uplus \ms_r \uplus \ms_f)) \land \npair[n-i]{(\reg',\ms')} \in \observations(W) \\
    & \qquad \Rightarrow \npair{(\reg,\ms \uplus \ms_r)} \in \observations(W)
  \end{align*}
\end{lemma}

@@

\begin{lemma}
  Say the instructions for \texttt{scall} $r([r_{j_1},\dots,r_{j_n}],[r_{k_1},\dots, r_{k_m}])$ are $i_0,\dots,i_l$, then
  \begin{align*}
    & m = \ms \uplus \ms_\stk \uplus \ms_f \\
    & \memSat{\ms \uplus \ms_\stk}{W}
    & \npair{\reg[\pc \mapsto c_{\var{scall}}][r_\stk \mapsto c_\stk]} \in \stdrr(W) \\
    & \ms_\stk = m | _{\start_\stk,\addrend_\stk} \\
    & c_\stk = ((\rwlx,\gl),\start_\stk,\addrend_\stk,\addr) \\
    & \start_\stk-1 \leq \addr \land \start_\stk \leq \addrend_\stk \land \addr \leq \addrend_\stk \\
    & c_{\var{scall}} = ((\perm,\gl)\start,\addrend,\addr) \\
    & \perm \in \{\exec, \rwx, \rwlx\} \\
    & \addrend - \addr \leq l  \\
    & \ms(\addr) = i_0, \dots, \ms(\addr+l) = \mathtt{jmp r}
    & (\reg[\pc \mapsto c_{\var{scall}}][r_\stk \mapsto c_\stk],m) \step[i'] (\reg_0[r_{k_1} \mapsto \reg(r_{k_1})]\dots[r_{k_m} \mapsto \reg(r_{k_m})][\pc \mapsto c_{\var{scall}}'][r_\stk \mapsto c_\stk'],m) \\
    & c_{\var{scall}}'.\mathrm{a} = \addr+l \\
    & \quad \Rightarrow \\
    & \qquad \exists \ms_\stk', \ms', \ms_r \ldotp \\
    & \qquad \quad W' = \revokeTemp{W}[\var{rn}_1\mapsto\iota^\sta(\temp, )][\var{rn}_2 \mapsto \iota^\pwl] \\
    & \qquad \quad \memSat[n-i']{\ms_\stk'\uplus \ms'}{W'}
  \end{align*}
\end{lemma}

>>>>

% Variables
\newcommand{\var}[1]{\mathit{#1}}
\newcommand{\hs}{\var{ms}}
\newcommand{\ms}{\hs}
\newcommand{\hv}{\var{hv}}
\newcommand{\rv}{\var{rv}}
\newcommand{\lv}{\var{lv}}
\newcommand{\gl}{\var{g}}
\newcommand{\pc}{\mathit{pc}}
\newcommand{\pcreg}{\mathrm{pc}}
\newcommand{\addr}{\var{a}}
\newcommand{\offset}{\var{offset}}
\newcommand{\word}{\var{w}}
\newcommand{\start}{\var{base}}
\newcommand{\addrend}{\var{end}}
\newcommand{\pwlv}{\var{pwl}}
\newcommand{\mem}{\var{mem}}
\newcommand{\reg}{\var{reg}}
\newcommand{\heapseg}{\var{ms}}
\newcommand{\heap}{\var{mem}}
\newcommand{\mode}{\var{mode}}
\newcommand{\perm}{\var{perm}}
\newcommand{\permp}{\var{permPair}}
\newcommand{\roll}{\var{roll}}
\newcommand{\instr}{\var{instr}}
\newcommand{\stdcap}[1][(\perm,\gl)]{\left(#1,\start,\addrend,\addr \right)}
\newcommand{\adv}{\var{adv}}
\newcommand{\msframe}{ms_\var{frame}}
\newcommand{\link}{\var{link}}
\newcommand{\stk}{\var{stk}}
\newcommand{\flag}{\var{flag}}
\newcommand{\nwl}{\var{nwl}}
\newcommand{\pwl}{\var{pwl}}
\newcommand{\sta}{\var{sta}}
\newcommand{\cnst}{\var{cnst}}
\newcommand{\olf}{\var{offsetLinkFlag}}
\newcommand{\prp}{\var{prp}}
\newcommand{\env}{\var{env}}
\newcommand{\cls}{\var{cls}}

@@

% Macros
\newcommand{\scall}[3]{\mathtt{scall} \; #1([#2],[#3])}

@@

\begin{proof}[Proof of Lemma~\ref{lem:correctness-f3}]
\lau{proof using scall lemma}
Assume the premises of the lemma.
Now define 
  \begin{align*}
    W = \;& [0 \mapsto \iota_{\malloc,0}] \\
          & [1 \mapsto \iota^\sta (\perma,\ms_\flag \uplus \ms_{f2})]\\
          & [2 \mapsto \iota^\sta (\perma,\ms_\link)]\\
          & [3 \mapsto \iota^{\nwl,p}_{\start_\adv,\addrend_\adv}]\\
  \end{align*}
Further define
\[
\ms' = \ms_\flag \uplus \ms_{f2} \uplus \ms_\link \uplus \ms_\adv
\]
If we can show 
\begin{equation}
\label{eq:f3-obs1}
  \npair[n+1]{(\reg\update{\pcreg}{c_{f3}}\update{r_\stk}{c_\stk},\ms_{\malloc} \uplus \ms' \uplus \ms_{\adv} \uplus \ms_\stk)} \in \observations(W),
\end{equation}
then using $\ms_{\var{frame}}$ as the frame and $m'$ as the resulting memory, we get that $m' = \ms'' \uplus \ms_r \uplus \ms_{\var{frame}}$ for some $\ms'$ and $\ms_r$ s.t. $\heapSat[1]{\ms''}{W}$. Region $1$ guarantees that the assertion flag is unchanged, so we have
\[
  m'(\flag) = 0
\]
So SFTS \ref{eq:f3-obs1}. To do so, we use Lemma \ref{lem:anti-red-obs}. Let $\ms_f$ be given. The execution proceeds as follows:
\[
  (\reg\update{\pcreg}{c_{f3}}\update{r_\stk}{c_\stk},\ms' \uplus \ms_\stk \uplus \ms_f) \step[i] (\reg',\ms' \uplus [\start_\stk \mapsto 1] \uplus \ms_\stk|_{\start_\stk+1,\addrend_\stk} \uplus \ms_f),
\]
where
\[
  \lookingat{(\reg',\ms')}{\scall{r}{}{}}{c_{\var{next}}}
\]
where $c_{\var{next}}$ is $c_{f3}$ adjusted to point to the next instruction, namely \texttt{pop r1}. Further we have \pointstostack{\reg'}{[\start_\stk \mapsto 1]}{\ms_\stk |_{\start_\stk+1,\addrend_\stk}} and $i$ is a suitable number of steps.

@@

            \npair[n_2]{(\reg_1\update{\pcreg}{\updatePcPerm{c_{g1}}},\ms_1)} \in \observations(W_1)
          %TODO elaborate on what happens here using anti reduction lemma and malloc spec.
          Use Lemma~\ref{lem:anti-red-obs}, so let $\ms_f'$ be given, then
            (\reg_1\update{\pcreg}{\updatePcPerm{c_{g1}}},\ms_1 \uplus \ms_f') \step[i] (\reg_2,\ms_2 \uplus \ms_\malloc' \uplus \ms_\env \uplus \ms_x \uplus \ms_\cls \uplus \ms_f')
          for some $i$. Where
          \item $\memSat[n_2-i]{\ms_2 \uplus \ms_\malloc'}{W_1\update{4}{\iota_\malloc}} $
          \item $\iota_\malloc \futurewk \iota_{\malloc,0}$
          \item $\ms_x = [x \mapsto 0]$
          \item $c_x = ((\rwx,\glob),x,x,x)$
          \item $\ms_\env = [\env \mapsto c_x]$
          \item $c_\env = ((\rwx,\glob,\env,\env,\env))$
          \item $c_{f4}$ is $c_{g1}$ updated to point at $f4$.
          \item $\ms_\cls$ contains $c_\env$, the $c_{f4}$, and the activation code.
          \item $c_\cls$ is a global enter capability for $\ms_\cls$ pointing to the activation code.
          \item $\reg_2 = reg_0\update{\pcreg}{\updatePcPerm{\reg_1(r_0)}}\update{r_0}{\reg_1(r_0)}\update{r_1}{c_\cls}$
          For world
            W_2 = W_1\update{4}{\iota_\malloc}\update{i_1}{\iota^\sta (\perm,\ms^\env \uplus \ms^\cls)}\iota{i_2}{\iota_x}
          where $i_1,i_2\not\in\dom(W_1)$ and $i_1 \neq i_2$ and $\iota_x$ is defined as
            \iota_x   & = (\perm, 0, \phi_\pub, \phi, H_x) \\
            \phi_\pub & = \{(0,1)\}^* \\
            \phi      & = (1,0) \union \phi_\pub \\
            H_x \; s \; \hat{W} & = \{\npair{\ms} \mid \ms(x) = s \land n > 0 \} \uplus \{\npair[0]{\ms}\}
          % TODO show $\iota_x$ is a region.
            \label{eq:g1-obs-to-show}
            \npair[n_2-i]{(\reg_2,\ms_2 \uplus \ms_\malloc' \uplus \ms_\env \uplus \ms_x \uplus \ms_\cls)} \in \observations(W_2)
          If $\reg_1(r_0).\perm \not\in \{\entry, \exec, \rwx, \rwlx\}$, then the execution fails after the jump and \label{eq:g1-obs-to-show} is thus trivially true.

          If $\reg_1(r_0).\perm \in \{\entry, \exec, \rwx, \rwlx\}$, then either $\execCond{}$ or $\entryCond{}$ holds for the capability in $\reg_1(r_0)$. Now use $W_2 \futurewk W_1$ with the approriate condition to get 
          \[
            \npair[n_2-i]{\updatePcPerm{\reg_1(r_0)}} \in \stder(W_2)
          \]
          which in turn gives us \ref{eq:g1-obs-to-show} if we can show the following
          \begin{enumproof}
            \item $\memSat[n_2-i]{\ms_2 \uplus \ms_\malloc' \uplus \ms_\env \uplus \ms_x \uplus \ms_\cls}{W_2}$
              % TODO follows from \ms_1 heap sat and the new regions.
            \item $\npair[n_2-i]{\reg_2} \in \stdrr$ \\
              Amounts to showing
              \begin{enumproof}
                \item $\npair[n_2-i]{\reg_2(r_0)} \in \stdvr(W_2)$ by assumption $\npair[n_2]{\reg_1} \in \stdrr(W_1)$ and $\stdvr$ monotonicity wrt. $\futurewk$
                \item $\npair[n_2-i]{c_\cls} \in \stdvr(W_2)$ \\
                  Let $n_3 < n_2 - i$ and $W_3 \futurestr W_2$ be given and show
                  \[
                    \npair[n_3]{\updatePcPerm{c_\cls}} \in \stder(W_3)
                  \]
                  To this and let $n_4 \leq n_3$, $\memSat[n_4]{\ms_3}{W_3}$, and $\npair[n_4]{\reg_3} \in \stdrr(W_3)$ and show
                  \begin{equation}
                    \label{eq:g1-c-cls-obs}
                    \npair[n_4]{(reg_3\update{\pcreg}{\updatePcPerm{c_\cls}},\ms_3)} \in \observations(W_3)
                  \end{equation}
                  Let $\ms_3^p$ and $\ms_3^t$ be memory segments such that $\ms_3 = \ms_3^p \uplus \ms_3^t$ and $\memSat[n_4]{\ms_3^p}{\revokeTemp{W}_3}$ (using Lemma~\ref{lem:priv-mono-like2})

                  Use Lemma~\ref{lem:anti-red-obs}, let $\ms_f''$ be given and take $\ms_r$ to be the part of $\ms_3^t$ that $\reg_3(r_\stk)$ does not govern. By the operational semantics, we know\footnote{the execution may fail, but then the confifuration is trivially in the observation relation.}
                  \[
                    (reg_3\update{\pcreg}{\updatePcPerm{c_\cls}},\ms_3^p \uplus \ms_3^t \uplus \ms_f'') \step[j] (\reg_4,\ms_4 \uplus \ms_3^t \uplus \ms_f'')
                  \]
                  where
                  \begin{itemize}
                  \item \lookingat{reg_4,\ms_4}{\scall{r_1}{}{r_0,r_1,r_\env}}{c_{\var{next}}}
                    \begin{itemize}
                    \item $c_{\var{next}}$ is the capability pointing to the next instruction.
                    \end{itemize}
                  \item \pointstostack{\reg_4}{\emptyset}{\ms_{\var{unused}}}
                    \begin{itemize}
                    \item \texttt{prepstack} did not fail, so the capability must be $\rwlx$ and follow the stack convention.
                    \end{itemize}
                  \item $\reg_3(r_0)$ is a $\glob$ capability.
                    \begin{itemize}
                    \item \texttt{reqglob} did not fail
                    \end{itemize}
                  \item $\ms_4(x) = 0$
                  \item $\reg_3(r_\env) = c_\env$
                  \end{itemize}
                  region $i_2$ (the $\iota_x$ region) can be in either state $0$ or $1$, so to make sure it is in state $0$, we use a private transition. So let $W_4$ be $\revokeTemp{W_3}$ with region $i_2$ in state $0$.
              \end{enumproof}
          \end{enumproof}
      \end{enumproof}
  \end{enumproof}

@@

\begin{lemma}[$\mathtt{scall}$ works]
  \label{lem:scall-works}
  If
  \begin{itemize}
  \item $\heapSat[\ms]{n}{\revokeTemp{W}}$ 
  \item $(\reg,\ms) \text{ is looking at }
    \mathtt{scall}(r(\overline{r_{\mathit{arg}}},
    \overline{r_{\mathit{priv}}})) \text{ followed by } c_{\mathit{next}}$
  \item $\reg \text{ points to stack with $\ms_\stk$ used and $\ms_{\mathit{unused}}$ unused}$
  \item[Hyp-Callee] If
    \begin{itemize}
    \item $\dom(\ms_{\mathit{unused}}) = \dom(\ms_{\mathit{act}}
      \uplus \ms_{\mathit{unused}}')$,
    \item $W' =
      \revokeTemp{W}[\iota^{\sta}(\temp,\ms_\stk\uplus\ms_{\mathit{act}}),\iota^{\pwl}(\dom(\ms_{\mathit{unused}}'))]$,
    \item $\memSat[n-1]{\ms''}{W'}$
    \item $\reg' \text{ points to stack with $\emptyset$ used and $\ms_{\mathit{unused}'}$ unused}$
    \item $\forall a \in \dom(\ms_{\var{unused}}') \ldotp \ms''(a) = 0$
    \item $\reg'= \reg_0[\pcreg\mapsto\updatePcPerm{\reg(r)},
      \overline{r_{\mathit{arg}}} \mapsto \reg(\overline{r_{\mathit{arg}}}),r_0
      \mapsto c_{\mathit{ret}}, r_\stk \mapsto \_, r \mapsto \reg(r)]$ 
    \item $(n,c_{\mathit{ret}}) \in \stdvr(W')$
    \end{itemize}

@@

    then we have that $(n', (\reg',\ms'' \uplus \ms_\stk \uplus \ms_{\var{unused}}'')) \in \observations(W'')$
  \end{itemize}
  Then 
  \begin{itemize}
    \item $(\reg,\ms \uplus \ms_\stk \uplus \ms_{\var{unused}}) \in \observations(W)$
  \end{itemize}

@@

\begin{lemma}[Anti-reduction for $\observations$]
\label{lem:anti-red-obs}
  \begin{align*}
    & \forall n,i,\reg, \reg', \ms, \ms', \ms_r, W, W' \ldotp\\
    & \quad (\forall \ms_f \ldotp (\reg,\ms \uplus \ms_r \uplus \ms_f) \step[i] (\reg',\ms' \uplus \ms_r \uplus \ms_f)) \land \npair[n-i]{(\reg',\ms')} \in \observations(W') \land W' \futurestr W \\
    & \qquad \Rightarrow \npair{(\reg,\ms \uplus \ms_r)} \in \observations(W)
  \end{align*}
\end{lemma}

@@
